--
title: "Introduction to slimmmingDive"
author: "Fer Arce"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


This is a minimal example to introduce the usage of slimmingDive with
southern elephant seals dive information kindly provided by the
Australian Integrated Marine Observing System (IMOS). It requires the external software JAGS (Just another Gibs
sampler) to be installed or slimmingDive will fail to install.

slimmingDive is not on cran, so oyou need to download and install it
locally. You can also use either `remotes` or `devtools` packages



We load the data and see what's on it:
```{r }
library(slimmingDive)

data(eles)
dim(eles)
```
it contains a large number of columns, innecesary to our purposes, so
we preprocess the data. This step is important as it will rename and
create few extra variables needed for further processing.

```{r }

Data <- formatDives(eles)

dim(Data)
str(Data)
```

Next step is to otain the order in which the inflection points are
generated. As the Drifing segment and drift rate depends on them

```{r }

Data[c('order','minresid')] <- t(apply(Data,1,RBSM,retrieve='both'))
Data$minresid <- as.numeric(Data$minresid)
```

That step will take around a couple of minutes. Next step is to
generate a large bunch of variables used in following steps

```{r }

D1 <- NewVarsVect(Data)
                                   # 30 secs


D1[c('NDE','ds')] <- t(apply(D1,1,NDE, extract='both'))

                                #6 seconds

```

we have generated all the variables for the candidate Drift dives,
let's see how they lloks:



```{r }


par(mfrow = c(2,2),mar = c(4,4,2,2))
lapply(split(D1, D1$ref), plotDrift, xlab = 'foraging days', ylab = 'Drift rate', xlim = c(-10, 245))

```
We have a bunch of dives with associated drift rates at any value

Next step would be to get ridden of those that are certainly no drift
dives:


```{r }
D1 <- driftFilter(D1)

par(mfrow = c(2,2),mar = c(4,4,2,2))
lapply(split(D1, D1$ref), plotDrift, xlab = 'foraging days', ylab = 'Drift rate', xlim = c(-10, 245))



```

A pattern aroses, so next step is to kalman. Because it is time
consuming, we can load data as if just kalmaned

```{r }
data(kalman)


D2 <- lapply(kalman, PostKalman)
D2 <- lapply(D2, function(fo) fo <- fo[fo$zetas > .99, ])


par(mfrow = c(2,2), mar = c(4,4,2,2))
lapply(D2, plotDrift, xlab = 'foraging days', ylab = 'Drift rate', xlim = c(-10, 245))

```
